%~~~~~~~~ Chapter ~~~~~~~~
\chapter{Introduction}

%
% Quantum mechanics --> technological advances --> limits --> quantum control
% --> possibilities -->
%

We are today on the verge of a technological revolution, a quantum leap\footnote{pun intended} that holds the promise for a revolution in chemistry, pharmacology, telecommunications, virtually any application in which optimization is a problem and even Condensed Matter Physics itself.
But let us start from the beginning.
\medbreak

The laws of \ac{qm} started developing at the beginning of the XX century. At that moment, it was commonly thought that only a few details in the laws of Physics remained unexplained, but as Physicists delved deeper in these unexplained phenomena (namely the photo-electric effect and the black body radiation), it became clear that there was some missing piece.

Step by step it was found that as the energy scales of the experiments grew smaller, the well-founded laws of \emph{classical} physics started failing.
% infrared catastrophe
% atomic model (electron radiation)
The development and understanding of the laws of \ac{qm} was fast and in merely 30 years there were entire research fields centered already in different aspects or applications of it.

Of particular interest to this thesis is the line of material science and condensed matter physics research. Prior to the understanding of \ac{qm}, there was no theory to explain the properties of any material.
Wether a piece of rock was able to conduct heat or electricity was essentially a matter of trial and error. Of course some experimental observations led Chemists to build a collection of somewhat empirical laws and group elements based on similar properties, but there was no microscopic understanding of the mechanisms leading to these properties.

The capabilities of \ac{qm} to predict electronic properties of different materials led to a revolution in science, technology and ultimately society. The extensive research on semiconductor materials during the 1960s was key to the invention of the transistor: a device that can block or allow the flow of electrical current at will.

The transistor is the basic building block for any electronic device. Nowadays there are millions if not billions of transistors in our computers or smartphones, but only a few thousand were enough to send 3 humans to the moon (and back!). For the last 50 years, transistors have operated based on the fine-tune of the band population of two neighbouring materials creating an effective potential barrier between them preventing electrons to move freely.
As the electronic industry developed the need for smaller transistors became apparent, the smaller the transistors are, the more can be fit in a particular device.
This race towards miniaturization pushed the development of new materials and techniques to size down transistors all the way to the current\footnote{In mass production, there are plenty of claims for smaller transistors.} $\sim\SI{14}{\nm}$. As expected, when we shrink the components of the transistors, we approach a limit in which quantum effects irrelevant until now may take a major role. The simplest obstacle that the shrinking of transistors have to deal with is the quantum tunnel effect, by which the electrons would be able to flow through the potential barrier, rendering the transistor unable to effectively stop electric current.

Quantum effects will play an important role as technology shrinks, but this may not be necessarily a problem.


\medbreak

Quantum limit is a feature. Feynman "control of atomic properties"

Quantum computers

Quantum simulations






%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
The development of Quantum Technologies has barely started and it already promises a revolution in many different fields, from medicine and pharmacology, computation, encryption, any application in which optimization is a problem and even the understanding of quantum mechanics itself.
There is probably nothing riskier than trying to make technological predictions, but even governments around the world are taking action to be in the forefront of this revolution\cite{QTF} and big companies such as Google, IBM or Microsoft are investing huge amounts of money in these topics. But let us start from the beginning.
\medbreak

% Quantum technologies
Nowadays we are barely getting a glimpse of the potential of quantum technologies to impact technology and science in general, but Quantum computation is not something that we just thought about.
Ever since we started the race to shrink every computer component, we headed towards the limit of quantum computation. A transistor is a device able to switch on and off an electrical current on command. They are the basic building block for any kind of computerized device. The first transistors were macroscopic, some centimeters and wired by hand. As the potential of computers grew, the need to create more, cheaper and smaller transistors launched the race for miniaturization. %Moore's Law is the empir
Nowadays a single transistor is the size of a few nanometers. At this scale  quantum effects start playing a significant role.
If there is a potential barrier stopping the electric current and it becomes smaller and smaller at some point the tunnel effect will be probable enough to render the transistor unable to stop any electric current at all.
This is one of the simplest limitations that computers face as we miniaturize components.

While it may seem that dealing with the inevitable quantum effects is a big inconvenience, scientists began to consider the possibility of using these effects to our benefit.
The first ideas about the possibilities for quantum computing started in the 80s and 90s\cite{Benioff1980, Feynman1982, Manin1980, Shor1994} showed that a quantum computer could tackle problems that a classical computer simply could not.

Since then there has been huge developments both theoretical and experimental towards the realization of a platform in which to perform \emph{quantum computations}.



\newpage
This thesis has somewhat been driven, at least in part, by a proposal dating all the way back to 1988\cite{Kane1988} on how to build a \emph{quantum computer}. In this proposal Bruce E. Kane suggested Phosphorus impurities in Silicon ($Si: \leftidx{^{31}}P$) as the building block for a silicon based nuclear spin quantum computer.
Any proposal for building quantum computers (or qubits, for that matter) has to deal with the conflict between two opposing requirements. On one hand we need the qubits to be well isolated from the environment so interactions with the outside world does not affect the quantum state of the computer. On the other hand, we need to be able to interact strongly and quickly with the qubits in order to change or read its state.

The idea to tackle this contradiction was to use the nuclear spin (1/2) of the $\leftidx{^{31}}P$ to store information, due to the relatively low concentration of nuclear spins in the system, and the electric cloud around them as mediators among $P$ impurities.
The way to control the interactions would be based on the application of electric gates which would distort the electronic states increasing or decreasing the hyperfine interaction (coupling between the nuclear and the electronic spin) as well as tuning the electronic configuration in between defects.
\medbreak

Ignoring the specific nature of the materials, the relevant ingredients in this proposal are a nuclear spin 1/2, which is expected to be isolated from the rest of the universe upon the application of an electric field, An electronic spin 1/2 located somewhat around the nuclear spin and with a big response to an electric field so the interaction between the two spins can be highly tuned.
\medbreak


% Graphene
In general, states near surfaces are very susceptible to external biases. In particular, \ac{2d} materials present highly manipulable electronic states since they only have the surface to live in.
Even before its discovery, graphene has draught a lot of attention. Many of its properties were expected well twenty years before its experimental discovery. While it is arguably true that there has been an over-hype bubble around its supposed miraculous applications, it is definitely true that graphene has been the gateway towards a whole new realm of phenomena in condensed matter physics. 

Initially most of its interest relied in the expectancy of a new Hall phase emerging without the need of an external magnetic field\cite{Kane2005a,Ostrovsky2008} and the experimental observation of the \ac{qhe}, a macroscopic effect of a purely quantum phenomenon\cite{Zhang2005}.
Also fact that some high-energy physics models were naturally realized in graphene was a huge motivation for this material to hoard the attention of a big part of the condensed matter community.

Certainly graphene has many interesting properties, but when two layers of graphene are placed one on top of each other there is a property that will be central for the development of this thesis.

Graphene bilayer is the only known material which can open a gap upon application of an electric field. It has been shown that an electri




%\newpage
%In order to start wrapping our minds around this Quantum Revolution, we need to look back to the beginning of the XIX century. Back then Physics was consider to be completely understood, except for a couple of seemingly simple experiments, namely the photoelectric effect and the black body radiation. These phenomena were the gateway to discover the realm of Quantum Mechanics.
%
%Many experiments followed and the theory was not left behind. Soon we had a framework with the mathematical tools to describe all these new phenomena. Nevertheless having the basic laws that rule this 
%
%
%\medbreak
%The world has seen all sorts of revolutions all throughout history. While not necessarily every revolution has turned the world into a better place, when we account just for the scientific or technological revolutions, this is usually the case.
%
%%At the end of the XVIII century, the invention of the steam engine and its later introduction in the industrial processes allowed a huge step forward for many societies and, for better or worse\cite{}%climate
%%, reshaped the world.
%%The Industrial Revolution changed the way we travel and work, it changed the scale at which humans interacted with the planet and, arguably, the universe.
%%The whole world has seen how virtually any index measuring quality of life has been raising ever since\cite{}.\footnote{Sadly this trend has not been homogeneous at all around the globe and these changes have been much more clear in some countries than others.}
%
%%Similarly, another huge revolution came along with
%The worldwide introduction of computers in most households and the construction of a worldwide web interconnecting ``everyone'' in the world has been one of the biggest revolutions over the last 30 years. The information technologies have changed the job-market landscape, the way humans learn, shop, the way we socialize even the way we fight wars. Even in their first stages computers were able to put two men on the moon and bring them back with less computational power than the cheapest smartphone available today.
%Our daily habits, our leisure time, even crucial historical events\cite{Alhindi2012} have been determined by the use of computers and the internet.
%
%For this revolution to happen, many agents had to be in place. The need for secure communication channels during World War II pushed the development of information theory and the later Cold War pushed scientists and engineers to build computers. But the bare necessity is not enough to develop the necessary technology. The first computers used \emph{macroscopic transistors} and \emph{macroscopic bits} when they were introduced. At first these essential pieces were handmade and manufactured one by one. It was not until the research in semiconductor physics developed microscopic versions of these pieces that the computers became cheap and easy to produce.
%\medbreak
%
%\newpage
%For nowadays computers to exist, it was necessary the development of Quantum Mechanics. Starting with the beginning of the XX century, physicist realised that Physics was not completely understood but rather there were many mysteries when 
%
%
%
%Naturally these revolutions did not just happen, it took many years of hard work and research to stablish the basis for these technologies to emerge. In particular, for computers to be created and be popularized it was necessary a deep understanding of the laws of quantum physics which led to the mastery of the semiconductor technologies.
%
%
%
%\newpage
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%The relation between research in fundamental Physics and technological innovations is not always easy to see. In the best of cases, it usually takes some years, at least a few decades, for fundamental research to find its way into the industry.
%While this is a completely fair question, it is quite short-sighted. Let us think of the latest revolution that changed our society.
%The introduction of computers in our everyday life and the access to a global network of communications was a revolution, maybe even comparable to the industrial revolution.
%
%We have had computers in most households for the last 30 years, broad access to the internet for over 20, yet, trying to pin-point the fundamental research that took us there is quite hard, if not impossible.
%We can think of the first transistor\footnote{actually there is a previous patent, from 1930,\cite{diode} but no hard proof that it was ever built or used.},
%the building block of every electronic device, which would take us to the 40's\cite{Ross}, but it would be unimaginable to reach that point without at least some understanding of the physics of semiconductors which had been studied for the previous 50 years. It would be insane trying to understand semiconductors without, at least, a few notions of quantum mechanics, electromagnetism...
%We could line up all the most brilliant minds of humanity that set the path towards today's technology: Landau, Bardeen, Bloch Dirac, Heisenberg, Bragg, Brillouin, Faraday, Gauss... countless others... Ask them about the application of their research, and not one of them would be able to describe a personal computer or the internet.
%
%That is why asking the researchers about the usefulness of their research is like asking newborn babies what will be their job position in 42 years time.
%
%Of course we have our own vision of what the research may bring upon the world, but let us keep in mind that this vision is nothing but a childish dream, since time and time again the applications have exceeded the wildest expectations of the researchers.\\
%
%
%At the end of the XIX century it was believed that Physics was solved. Today we know it is not. In particular, Condensed Matter Physics is an area in which even when we know the physical laws governing every interaction there is still much that we do not understand. Quantum Mechanics gives us the tools to describe any problem that we want to describe. From superconductors to topological phases or spin liquids, the Hamiltonian is known and yet we cannot solve it. The last fifty or sixty years of condensed matter physicist have been devoted to approximations and simplifications that would allow us to strip the problems to its basic components in an attempt to understand the microscopical mechanism behind so many interesting phenomena.
%The thorough investigation of a variety of problems while not completely successful in every area, has had a great side effect. The cumulative experience of decades of condensed matter physics research has granted us the ability to control matter with atomic precision in a wide range of situations. The fulfillment of such a dream\cite{Feynman1982}
%
%Quantum technologies\cite{QTF}
%
%The research on semiconductor physics can be hard to pin-point, 
%
%
%In this work we study a variety of properties regarding defected graphene bilayer. Starting from the basic concepts of graphene, we build our way to the role of sp$^3$-defects in graphene bilayer such as chemisorbed Hydrogen adatoms.
%First we study the properties of these defects one by one in isolated scenarios. Then we wonder the effects of neighboring defects
