\chapter{Publications}
The following is a complete list of my publications:

\begin{itemize}
   \item N. A. García-Martínez, B. Valenzuela, S. Ciuchi, E. Cappelluti, M. J. Calderon, and E. Bascones, ``Coupling of the as a 1g phonon to magnetism in iron pnictides'', Physical Review B, vol. 88, no. 16, p. 165106, 2013.\onlinecite{Garcia-Martinez2013}

   \item N. A. García-Martínez, A. G. Grushin, T. Neupert, B. Valenzuela, and E. V. Castro, ``Interaction-driven phases in the half-filled spinless honeycomb lattice from exact diagonalization'', Physical Review B, vol. 88, p. 245123, dec 2013.\onlinecite{Garcia-Martinez2013a}

   \item N. A. García-Martínez, J. L. Lado, and J. Fernández-Rossier, ``Quantum spin hall phase in multilayer graphene'', Physical Review B, vol. 91, p. 235451, jun 2015.\onlinecite{Garcia-Martinez2015}

   \item J. Lado, N. García-Martínez, and J. Fernández-Rossier, ``Edge states in
graphene-like systems'', Synthetic Metals, vol. 210, pp. 56–67, 2015.\onlinecite{Lado2015}

   \item N. A. García-Martínez, J. L. Lado, D. Jacob, and J. Fernández-Rossier, ``Anomalous magnetism in hydrogenated graphene'', Phys. Rev. B, vol. 96, p. 024403, 2017.\onlinecite{Garcia-Martinez2017}

   \item D. Carvalho, N. A. García-Martínez, J. L. Lado, and J. Fernández-Rossier, ``Real-space mapping of topological invariants using artificial neural networks'', Physical Review B, vol. 97, no. 11, p. 115453, 2018.\onlinecite{Carvalho2018}

   \item R. Ortiz, N. A. García-Martínez, J. L. Lado, and J. Fernández-Rossier, ``Electrical spin manipulation in graphene nanostructures'', Physical Review B, vol. 97, no. 19, p. 195425, 2018.\onlinecite{Ortiz2018}

   \item C. Cardoso, D. Soriano, N. A. García-Martínez, and J. Fernández-Rossier, ``Van der waals spin valves'', Phys. Rev. Lett., vol. 121, no. 6, p. 67701, 2018.\onlinecite{Cardoso2018}

   \item R. Ortiz, R. Boto, N. García-Martínez, J. Sancho-García, M. Melle-Franco, and J. Fernández-Rossier, ``Exchange rules for diradical $\pi$-conjugated hydrocarbons'', Nano Lett., vol. 19, no. 9, 2019.\onlinecite{Ortiz2019}

   \item N. A. García-Martínez and J. Fernández-Rossier, ``Designer fermion models in functionalized graphene bilayers'', Phys. Rev. Res., vol. 1, no. 3, pp. 1–5, 2019.\onlinecite{Garcia-Martinez2019}
\end{itemize}
